Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-04-19 Manuscript/manuscript.Rmd
manuscript edits

## VARIABLE IMPORTANCES FROM TOP PERFORMING MODELS CONFIRMED BIOLOGICAL FUNCTIONALITY OF KNOWN GENOMIC ANNOTATIONS

@@

Given the prevalence of TFBS as the most consistently top ranked predictive features across resolution and cell line, we ventured to further compare performances of annotation specific models. After applying SMOTE re-sampling with distance type predictors, random forest models built on sets of histone modifications, chromatin state segmentations, and TFBS were compared to a model that included all three types together. For the GM12878 cell line, models built simply on TFBS were found to perform just as well as models that included all 3 sets of annotations according to F1-Score and MCC, consistently across resolution (Figure 6). Similar results were found when evaluating the K562 cell line (supplementary Figure 11).

>>>>

## VARIABLE IMPORTANCES FROM TOP PERFORMING MODELS CONFIRMED BIOLOGICAL FUNCTIONALITY OF KNOWN TRANSCRIPTION FACTOR BINDING SITES

@@

Given the prevalence of TFBS as the most consistently top ranked predictive features across resolution and cell line, we ventured to further compare performances of annotation specific models. After applying SMOTE re-sampling with distance type predictors, random forest models built on sets of histone modifications, chromatin state segmentations, and TFBS were compared to a model that included all three types together. For the GM12878 cell line, models built simply on TFBS were found to perform just as well as models that included all 3 sets of annotations according to F1-Score and MCC, consistently across resolution (Figure 6). Similar results were found when evaluating the K562 cell line (supplementary Figure 11). For both resolutions, models built on only histone modifications were found to have the lowest F1-Scores and MCC, across resolutions, followed by models built on chromatin state segmentations.

Distance profiles on a $log_{2}$ scale were found to exhibit significantly different densities between classes of genomic bins for TFBS with highly ranked predictive importances. Using permutation tests we found that the transcription factors SMC3a, ZNF143, CTCF, and RAD21 tended to have binding peaks that that were located closer to the centers of 10 kb sized genomic bins that overlapped with flanked TAD boundaries than to bins that did not overlap with flanked TAD boundaries across both cell lines (Figure 7 A-H). Similar distance profiles for the same TFBS were obtained when assessed at lower resolutions, across both cell lines (Supplementary Figures X-X)
